---
title: "Quality Control"
author: "Maia Smith"
date: "March 24, 2015"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
opts_chunk$set(out.width='1000px', dpi=200)
```

Quality Control
=============

Note: To get the hover information working correctly on Plotly graphs (the interactive graphs), do the following: 
1. Hover anywhere over graph
2. You'll see icons appear at the top right-hand corner of the graph
3. Select the icon for "Show closest data on hover"

Compare controls and treatments
-----------------------------------------------
```{r, eval=FALSE, echo=FALSE}
plot <- ggplot(sytoxG_data, aes(x=as.numeric(time_elapsed), y=as.numeric(phenotype_value), text=Compound, group=Compound)) +
  geom_line() +
  xlab("Time Elapsed") +
  ylab("Sytox Green") +
  ggtitle("Sytox Green Over Time - Control vs Treatment") +
  facet_grid(~empty, scales = "fixed") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white"), 
        axis.text = element_blank())
response <- py$ggplotly(plot, kwargs=list(world_readable=FALSE, filename="SG_control_vs_treatment", fileopt="overwrite"))
url <- response$response$url
plotly_iframe <- paste("<center><iframe scrolling='no' seamless='seamless' style='border:none' src='", url, 
    "/800/1200' width='800' height='800'></iframe><center>", sep = "")
```
`r I(plotly_iframe)`

```{r, eval=FALSE, echo=FALSE}
plot <- ggplot(confluency_data, aes(x=as.numeric(time_elapsed), y=as.numeric(phenotype_value), text=Compound, group=Compound)) +
  geom_line() +
  xlab("Time Elapsed") +
  ylab("Confluency") +
  ggtitle("Confluency Over Time - Control vs Treatment") +
  facet_grid(~empty, scales = "fixed") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white"), 
        axis.text = element_blank())
response <- py$ggplotly(plot, kwargs=list(world_readable=FALSE, filename="Con_control_vs_treatment", fileopt="overwrite"))
url <- response$response$url
plotly_iframe <- paste("<center><iframe scrolling='no' seamless='seamless' style='border:none' src='", url, 
    "/800/1200' width='800' height='800'></iframe><center>", sep = "")
```
`r I(plotly_iframe)`

Compare plates for controls and treatments
-----------------------------------------------

```{r, echo=FALSE}
# compare sytoxG sparklines, faceted by plates, empty
ggplot(sytoxG_data, 
       aes(x=as.numeric(time_elapsed), y=as.numeric(phenotype_value), group=Compound)) +
  geom_line() +
  xlab("Time Elapsed") +
  ylab("Sytox Green") +
  ggtitle("Sytox Green Over Time - By Plate") +
  facet_grid(empty ~ Plate, scales = "fixed") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white"))

# compare confluency sparklines, faceted by plates, empty
ggplot(confluency_data, 
       aes(x=as.numeric(time_elapsed), y=as.numeric(phenotype_value), group=Compound)) +
  geom_line() +
  xlab("Time Elapsed") +
  ylab("Confluency") +
  ggtitle("Confluency Over Time - By Plate") +
  facet_grid(empty ~ Plate, scales = "fixed") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white"))
```

Compare sparklines for each plate, controls vs treatments, with mean and sd for each plate and control/treatment
-----------------------------------------------

```{r, echo=FALSE}
#sytoxG - calculate mean and sd values for each plate, empty vs not empty
sytoxG_mean_sd_empty <- ddply(sytoxG_data, ~ Plate * time_elapsed * empty, summarize,
                              mean = mean(phenotype_value), sd = sd(phenotype_value))

#sytoxG - plot mean and sd values for each plate, neg.control vs others - using geom_ribbon()
ggplot(sytoxG_mean_sd_empty, 
       aes(x=as.numeric(time_elapsed), y=as.numeric(mean), colour = as.factor(Plate), group = Plate)) +
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), alpha=0.2) + 
  facet_grid(~ empty, scales = "fixed") +
  xlab("Time Elapsed") +
  ylab("Sytox Green Mean & SD") +
  scale_colour_discrete(name="Plate") +
  ggtitle("Sytox Green\nPlate Mean & SD (line size)\nNegative Control T/F (facets)") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white")) 

#confluency - calculate mean and sd values for each plate, empty vs not empty
con_mean_sd_empty <- ddply(confluency_data, ~ Plate * time_elapsed * empty, summarize,
                              mean = mean(phenotype_value), sd = sd(phenotype_value))

#confluency - plot mean and sd values for each plate, neg.control vs others - using geom_ribbon()
ggplot(con_mean_sd_empty, 
       aes(x=as.numeric(time_elapsed), y=as.numeric(mean), colour = as.factor(Plate), group = Plate)) +
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), alpha=0.2) + 
  facet_grid(~ empty, scales = "fixed") +
  xlab("Time Elapsed") +
  ylab("Confluency Mean & SD") +
  scale_colour_discrete(name="Plate") +
  ggtitle("Confluency\nPlate Mean & SD (line size)\nNegative Control T/F (facets)") +
  theme(panel.grid = element_blank(),
        axis.ticks.length = unit(0, "cm"),
        panel.background = element_rect(fill = "white")) 
```

